---
title: "Visualizaation"
subtitle: "Part II"
output: github_document
---


```{r setup}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() +  theme(legend.position = "bottom"))

options(
  ggplots2.continuous.color = "viridis",
  ggplots2.continuous.fill = "viridus"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
# Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>% 
  select(name, id, everything())

weather_df
```

## `group_by`
```{r}
weather_df %>% 
  group_by(name, month)
  ## %>% ungroup(month)
```

## counting things
count month /name observations
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs=n())
```

we can use `count`
```{r}
weather_df %>% 
  count(month, name = "n_obs")
```

**Never** use base R's `table`
```{r eval=FALSE}
weather_df %>% 
  pull(month) %>% 
  table() ##bad idea
```

other helpful counters
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs=n(),
    n_days = n_distinct(date))
```

## A digression on 2x2 tables
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ "")
    ) %>% 
  janitor::tabyl(name, cold)

```

## General summaries

you can do plots of summaries.
```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    meadian_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe! So we can make a plot.
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    meadian_tmin = median(tmin, na.rm = TRUE)
    ) %>% 
      ggplot(aes(x = month, y = mean_tmax, color = name)) +
      geom_point() +
      geom_line()
```

Suppose you want to summarize many columns.
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```

Reminder: sometimes your results are easier to read in another format.
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1)
```

